---
title: "project"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data Preprocessing  


```{r, cache = T}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```

##Download the Data
```{r, cache = T}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="curl")
}
```  

##Read the Data
read the two csv files into two data frames:
```{r, cache = T}
trainRaw <- read.csv("./data/pml-training.csv")
testRaw <- read.csv("./data/pml-testing.csv")
dim(trainRaw)
dim(testRaw)
```
the training data set contains 19622 observations and 160 variables.
the testing data set contains 20 observations and 160 variables. 
the "classe" variable in the training set is the outcome to predict. 

## Clean the data
clean the data and get rid of observations with missing values as well as some meaningless variables:
```{r, cache = T}
sum(complete.cases(trainRaw))
```

remove columns that contain missing values:
```{r, cache = T}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
```  

remove some columns that do not contribute much to the accelerometer measurements:
```{r, cache = T}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
dim(testCleaned)
dim(trainCleaned)
```
the cleaned training data set contains 19622 observations and 53 variables, while the testing data set contains 20 observations and 53 variables. The "classe" variable is still in the cleaned training set.

## Slice the data
plit the cleaned training set into a pure training data set (70%) and a validation data set (30%). 
use the validation data set to conduct cross validation in future steps: 
```{r, cache = T}
set.seed(1905) # For reproducibile purpose
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Data Modeling
fit a predictive model for activity recognition using Random Forest algorithm because it automatically selects important variables and is robust to correlated covariates & outliers in general. 
use 5-fold cross validation when applying the algorithm:
```{r, cache = T}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```

estimate the performance of the model on the validation data set:
```{r, cache = T}
predictRf <- predict(modelRf, testData)
confusionMatrix(as.factor(testData$classe), predictRf)
```
the estimated accuracy of the model is 99.1% 

## Predicting for Test Data Set
apply the model to the original testing data set downloaded from the data source. We remove the `problem_id` column first.  
```{r, cache = T}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  


Decision Tree Visualization
```{r, cache = T}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel) # fast plot
```