---
title: "Course Project: Shiny Application and Reproducible Pitch"
author: "yuxi"
date: "2/13/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The World's Telephones

- This project introduces the number of telephones in various regions of the world (in thousands).
- Format: A matrix with 7 rows and 8 columns. The columns of the matrix give the figures for a given region, and the rows the figures for a year.

## Data Format

- The regions are: North America, Europe, Asia, South America, Oceania, Africa, Central America.
- The years are: 1951, 1956, 1957, 1958, 1959, 1960, 1961.

## Data Summary

```{r cars, echo = FALSE}
library(tidyr)   # For pivoting the data into long form 
library(tibble)  # For converting the rownames (Year) to a column
library(scales)  # For scaing the y-axis and labels
library(ggplot2) # For the plot
summary(WorldPhones)
```

## Slide with Plot

```{r pressure}
WorldPhones %>%
  as.data.frame() %>%
  rownames_to_column("Year") %>%
  pivot_longer(cols=-Year, names_to="Country", values_to="Users") %>%
  ggplot(aes(Year, Users, group=Country, col=Country)) +
  geom_line() +
  scale_y_log10(n.breaks=5, labels = trans_format("log10", math_format(10^.x))) +
  theme_minimal() 
```

